---
title: "ANCOVA Test Using R Markdown"
#author: "Jeffrey Snowden"
#date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Jeffrey/Desktop/R")
library(tidyverse)
library(ggpubr)
library(rstatix)
library(broom)
library(datarium)
library(haven)
library(foreign)
library(ggplot2)
library(emmeans)
library(gridExtra)
library(cowplot)
library(pander)

setwd("C:/Users/Jeffrey/Desktop/R")

urbanhs <- read_sav("UrbanHS.sav")

urbanhs$NGA <- as.factor(urbanhs$NGA)

urbanhs <- urbanhs %>%
  select(ID, NGA, Grade9GPA, MeanReadMath, RatioCreditsAttemptedEarned, MinorityWhite, FRLYesNo) %>%
  rename(MeanRM = MeanReadMath, RatioCredits = RatioCreditsAttemptedEarned)
```

This R Markdown follows the guide Joe posted: [ANCOVA in R](https://www.datanovia.com/en/lessons/ancova-in-r/). I have not checked all of the assumptions, yet.

## Assumptions

These plots check the linearity between each covariate and the outcome variable. GPA and credit ratio appear to have a mostly linear relationship, while GPA and the mean of the TCAP Math and Reading scores is not linear as they cross. I'm not sure how much of a problem this is.

```{r echo=FALSE, warning=FALSE}
#Create scatter plot for covariate MeanRM and GPA
plot1 <- ggscatter(
  urbanhs, x = "MeanRM", y = "Grade9GPA",
  color = "NGA", add = "reg.line"
  )+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"), color = NGA)
  )+
  xlab("Mean Reading and Math Score")+
  ylab("9th Grade GPA")+
  ggtitle("9th Grade GPA and TCAP Scores")

plot1

#Create scatter plot for covariate RatioCredits and GPA
plot2 <- ggscatter(
  urbanhs, x = "RatioCredits", y = "Grade9GPA",
  color = "NGA", add = "reg.line"
)+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"), color = NGA)
  )+
  xlab("Credits Earned to Credits Attempted")+
  ylab("9th Grade GPA")+
  ggtitle("9th Grade GPA and Credit Ratio")

plot2
```

## The ANCOVA Model
```{r echo=FALSE, warning=FALSE, message=FALSE}
#Run the ANCOVA controlling for MeanRM and RatioCredits
#model <- aov(lm(Grade9GPA ~ MeanRM + RatioCredits + NGA, data = urbanhs), type="III")
#pander(summary(model), style = "rmarkdown")
```

I ran an ANCOVA model with the same covariates, the output matches what I got from SPSS. I can't figure out how to get the sum of squares with this function.
```{r echo=FALSE, include=FALSE}
res.aov <- urbanhs %>% anova_test(Grade9GPA ~ RatioCredits + MeanRM + NGA, type=3)
```

```{r echo=FALSE}
pander(get_anova_table(res.aov), style = "rmarkdown", split.tables = 120)
```

## Estimated Marginal Means
The estimated marignal means below match the output I got from SPSS when using the Bonferroni adjustment.
```{r echo=FALSE, include=FALSE}
#Get the Estimated Marginal Means
pwc <- urbanhs %>%
  emmeans_test (
    Grade9GPA ~ NGA, covariate = c("RatioCredits", "MeanRM"),
    p.adjust.method = "bonferroni"
  )
#pander(pwc, style = "rmarkdown")
```

```{r echo=FALSE}
#Print the EMM's
pander(get_emmeans(pwc), style = "rmarkdown", split.table = 120)

# Visualization: line plots with p-values
pwc <- pwc %>% add_xy_position(x = "NGA", fun = "mean_se")
ggline(get_emmeans(pwc), x = "NGA", y = "emmean") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) + 
  labs(
    subtitle = get_test_label(res.aov, detailed = TRUE),
    caption = get_pwc_label(pwc)
  )+
  #xlab("9th Grade Academy")+
  ylab("Estimated Marginal Means")+
  scale_x_discrete("9th Grade Academy", labels = c("Did not Attend", "Attended"))+
  ggtitle("Estimate Marginal Means for 9th Grade Academy")
```
